---
title: "Project"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Read Data
```{r}
set.seed(1)
TD = read.csv("train_data.csv")

#Split training data into 80-20 train/validation split.
train_ind = sample(length(TD$Status), 0.70*length(TD$Status))
train = TD[train_ind,]
valid = TD[-train_ind,]
test = read.csv("test_data.csv")
```

## Test With Best Model
```{r}
arr = c(4,1,5,1,5)
lda.fit = suppressWarnings(  lda(factor(Status) ~  poly(smoking,arr[1],raw=T) + poly(cholesterol,arr[2],raw=T) + poly(behavior,arr[3],raw=T) +  poly(age,arr[4],raw=T) + poly(assay,arr[5],raw=T), data=train) )
lda.pred = predict(lda.fit, test)$class
IDs = 1001:(1001+length(test$BP)-1)
results = ifelse(lda.pred==1,TRUE,FALSE)
toExport = data.frame("ID"=IDs,"Category"=results)
write.csv(toExport,file="testPredictions.csv",row.names=FALSE)
```


## Best Subset
```{r}
library(leaps)
subset.fit = regsubsets(Status~.,data=train,nvmax=length(train))
subset.sum = summary(subset.fit)
subset.sum
par(mfrow=c(2,2)) 
plot(subset.sum$rss, xlab="Num of Vars", ylab="RSS", type="l")
points(which.min(subset.sum$rss), subset.sum$rss[which.min(subset.sum$rss)], col="red",cex=2,pch=20)
plot(subset.sum$adjr2, xlab="Num of Vars", ylab="Adjusted RSq", type="l") 
points(which.max(subset.sum$adjr2), subset.sum$adjr2[which.max (subset.sum$adjr2)], col="red",cex=2,pch=20) 
plot(subset.sum$cp, xlab ="Num of Vars",ylab="Cp",type="l") 
points (which.min(subset.sum$cp), subset.sum$cp[which.min (subset.sum$cp)], col="red",cex=2,pch=20) 
plot(subset.sum$bic, xlab="Num of Vars",ylab="BIC", type="l") 
points (which.min(subset.sum$bic), subset.sum$bic[which.min(subset.sum$bic)],col="red",cex=2,pch=20)
```

Which predictors to use according to Adjusted RSq
```{r}
subset.coeff = coef(subset.fit, which.max(subset.sum$adjr2))
```

## LogRed/LDA/QDA for Best Subset

Perform LogReg with chosen predictors
```{r}
bestVal = 0
for (x in 1:500) {
  arr = sample(1:5, 6, replace=TRUE)
  glm.fit = suppressWarnings(  glm(factor(Status) ~  poly(smoking,arr[1],raw=T) + poly(cholesterol,arr[2],raw=T) + poly(behavior,arr[3],raw=T) +  poly(age,arr[4],raw=T) + poly(gold_standard,arr[5],raw=T) + poly(assay,arr[6],raw=T), data=train, family="binomial") )
  glm.prob = predict(glm.fit, valid, type="response")
  glm.pred=rep(0,length(valid$Status))
  glm.pred[glm.prob>.5]=1
  if (mean(glm.pred==valid$Status) > bestVal) {
    glm.bestProb = glm.prob
    bestArr = arr
    bestVal = mean(glm.pred==valid$Status)
  }
}
bestArr
bestVal
```

ROC Curve with chosen predictors
```{r}
library(ROCR)
pred = prediction(glm.bestProb, valid$Status)
perf = performance(pred, "tpr", "fpr")
plot(perf, main="ROC Curve")
```

Perform LDA with chosen predictors
```{r}
library(MASS)
bestVal=0
for (x in 1:500) {
  arr = sample(1:5, 6, replace=TRUE)
  lda.fit = suppressWarnings(  lda(factor(Status) ~  poly(smoking,arr[1],raw=T) + poly(cholesterol,arr[2],raw=T) + poly(behavior,arr[3],raw=T) +  poly(age,arr[4],raw=T) + poly(gold_standard,arr[5],raw=T) + poly(assay,arr[6],raw=T), data=train) )
lda.pred = predict(lda.fit, valid)$class
  if (mean(lda.pred==valid$Status) > bestVal) {
    bestArr = arr
    bestVal = mean(lda.pred==valid$Status)
  }
}
bestArr
bestVal
```

Perform QDA with chosen predictors
```{r}
bestVal = 0
for (x in 1:500) {
  arr = sample(1:5, 6, replace=TRUE)
  qda.fit = suppressWarnings(  qda(factor(Status) ~  poly(smoking,arr[1],raw=T) + poly(cholesterol,arr[2],raw=T) + poly(behavior,arr[3],raw=T) +  poly(age,arr[4],raw=T) + poly(gold_standard,arr[5],raw=T) + poly(assay,arr[6],raw=T), data=train) )
qda.pred = predict(qda.fit, valid)$class
  if (mean(qda.pred==valid$Status) > bestVal) {
    bestArr = arr
    bestVal = mean(qda.pred==valid$Status)
  }
}
bestArr
bestVal
```

We see from LogReg summary some of the predictors are not significant. We now run a model without them.

LogReg with less predictors
```{r}
bestVal = 0
for (x in 1:500) {
  arr = sample(1:5, 5, replace=TRUE)
  glm.fit = suppressWarnings(  glm(factor(Status) ~  poly(smoking,arr[1],raw=T) + poly(cholesterol,arr[2],raw=T) + poly(behavior,arr[3],raw=T) +  poly(age,arr[4],raw=T) + poly(assay,arr[5],raw=T), data=train, family="binomial") )
  glm.prob = predict(glm.fit, valid, type="response")
  glm.pred=rep(0,length(valid$Status))
  glm.pred[glm.prob>.5]=1
  if (mean(glm.pred==valid$Status) > bestVal) {
    glm.bestProb = glm.prob
    bestArr = arr
    bestVal = mean(glm.pred==valid$Status)
  }
}
bestArr
bestVal
```

ROC Curve with chosen predictors
```{r}
library(ROCR)
pred = prediction(glm.prob, valid$Status)
perf = performance(pred, "tpr", "fpr")
plot(perf,main="ROC Curve")
```

LDA with less predictors
```{r}
bestVal=0
for (x in 1:500) {
  arr = sample(1:5, 5, replace=TRUE)
  lda.fit = suppressWarnings(  lda(factor(Status) ~  poly(smoking,arr[1],raw=T) + poly(cholesterol,arr[2],raw=T) + poly(behavior,arr[3],raw=T) +  poly(age,arr[4],raw=T) + poly(assay,arr[5],raw=T), data=train) )
lda.pred = predict(lda.fit, valid)$class
  if (mean(lda.pred==valid$Status) > bestVal) {
    bestArr = arr
    bestVal = mean(lda.pred==valid$Status)
  }
}
bestArr
bestVal
```

QDA with less predictors
```{r}
bestVal = 0
for (x in 1:500) {
  arr = sample(1:5, 5, replace=TRUE)
  qda.fit = suppressWarnings(  qda(factor(Status) ~  poly(smoking,arr[1],raw=T) + poly(cholesterol,arr[2],raw=T) + poly(behavior,arr[3],raw=T) +  poly(age,arr[4],raw=T) + poly(assay,arr[5],raw=T), data=train) )
qda.pred = predict(qda.fit, valid)$class
  if (mean(qda.pred==valid$Status) > bestVal) {
    bestArr = arr
    bestVal = mean(qda.pred==valid$Status)
  }
}
bestArr
bestVal
```

## LASSO
Perform Lasso to see if coefficients approach zero
```{r}
library (glmnet)
grid =10^seq(0,-4,length=100)
X = model.matrix(train$Status ~ . -1, data=train)[,-11]
Y = train$Status
lasso.ck=glmnet(X,as.factor(Y),alpha=1,lambda=grid,family="binomial")
plot(lasso.ck, xvar="lambda",main="Lasso Regression")
```

Perform Lasso CV to find best lambda
```{r}
set.seed(4)
lasso.cv =cv.glmnet(X,as.factor(Y),alpha=1,lambda=grid,family="binomial")
plot(lasso.cv, main="lambda CV")
bestlam =lasso.cv$lambda.min
```
See which coefficients approach 0
```{r}
lasso.zr=glmnet(X,Y,alpha=1,lambda=grid)
lasso.coef=predict(lasso.zr,type="coefficients",s=bestlam)[1:11,]
lasso.coef
```

## LogRed/LDA/QDA for Lasso

Perform LogReg with chosen predictors
```{r}
bestVal = 0
for (x in 1:500) {
  arr = sample(1:5, 7, replace=TRUE)
  glm.fit = suppressWarnings(  glm(factor(Status) ~  poly(smoking,arr[1],raw=T) + poly(cholesterol,arr[2],raw=T) + poly(behavior,arr[3],raw=T) + poly(BMI,arr[4],raw=T) +  poly(age,arr[5],raw=T) + poly(gold_standard,arr[6],raw=T) + poly(assay,arr[7],raw=T), data=train, family="binomial") )
  glm.prob = predict(glm.fit, valid, type="response")
  glm.pred=rep(0,length(valid$Status))
  glm.pred[glm.prob>.5]=1
  if (mean(glm.pred==valid$Status) > bestVal) {
    glm.bestProb = glm.prob
    bestArr = arr
    bestVal = mean(glm.pred==valid$Status)
  }
}
bestArr
bestVal
```

ROC Curve with chosen predictors
```{r}
library(ROCR)
pred = prediction(glm.bestProb, valid$Status)
perf = performance(pred, "tpr", "fpr")
plot(perf, main="ROC Curve")
```

Perform LDA with chosen predictors
```{r}
bestVal=0
for (x in 1:500) {
  arr = sample(1:5, 7, replace=TRUE)
  lda.fit = suppressWarnings(  lda(factor(Status) ~  poly(smoking,arr[1],raw=T) + poly(cholesterol,arr[2],raw=T) + poly(behavior,arr[3],raw=T) + poly(BMI,arr[4],raw=T) +  poly(age,arr[5],raw=T) + poly(gold_standard,arr[6],raw=T) + poly(assay,arr[7],raw=T), data=train) )
lda.pred = predict(lda.fit, valid)$class
  if (mean(lda.pred==valid$Status) > bestVal) {
    bestArr = arr
    bestVal = mean(lda.pred==valid$Status)
  }
}
bestArr
bestVal
```

Perform QDA with chosen predictors
```{r}
bestVal = 0
for (x in 1:500) {
  arr = sample(1:5, 7, replace=TRUE)
  qda.fit = suppressWarnings(  qda(factor(Status) ~  poly(smoking,arr[1],raw=T) + poly(cholesterol,arr[2],raw=T) + poly(behavior,arr[3],raw=T) + poly(BMI,arr[4],raw=T) +  poly(age,arr[5],raw=T) + poly(gold_standard,arr[6],raw=T) + poly(assay,arr[7],raw=T), data=train) )
qda.pred = predict(qda.fit, valid)$class
  if (mean(qda.pred==valid$Status) > bestVal) {
    bestArr = arr
    bestVal = mean(qda.pred==valid$Status)
  }
}
bestArr
bestVal
```

We see from LogReg summary some of the predictors are not significant. We now run a model without them.

LogReg with less predictors
```{r}
bestVal = 0
for (x in 1:500) {
  arr = sample(1:5, 4, replace=TRUE)
  glm.fit = suppressWarnings(  glm(factor(Status) ~  poly(smoking,arr[1],raw=T) + poly(cholesterol,arr[2],raw=T) +   poly(age,arr[3],raw=T) + poly(assay,arr[4],raw=T), data=train, family="binomial") )
  glm.prob = predict(glm.fit, valid, type="response")
  glm.pred=rep(0,length(valid$Status))
  glm.pred[glm.prob>.5]=1
  if (mean(glm.pred==valid$Status) > bestVal) {
    glm.bestProb = glm.prob
    bestArr = arr
    bestVal = mean(glm.pred==valid$Status)
  }
}
bestArr
bestVal
```

ROC Curve with chosen predictors
```{r}
library(ROCR)
pred = prediction(glm.bestProb, valid$Status)
perf = performance(pred, "tpr", "fpr")
plot(perf,main="ROC Curve")
```

LDA with less predictors
```{r}
bestVal=0
for (x in 1:500) {
  arr = sample(1:5, 4, replace=TRUE)
  lda.fit = suppressWarnings(  lda(factor(Status) ~  poly(smoking,arr[1],raw=T) + poly(cholesterol,arr[2],raw=T) +  poly(age,arr[3],raw=T) + poly(assay,arr[4],raw=T), data=train) )
lda.pred = predict(lda.fit, valid)$class
  if (mean(lda.pred==valid$Status) > bestVal) {
    bestArr = arr
    bestVal = mean(lda.pred==valid$Status)
  }
}
bestArr
bestVal
```

QDA with less predictors
```{r}
bestVal = 0
for (x in 1:500) {
  arr = sample(1:5, 4, replace=TRUE)
  qda.fit = suppressWarnings( qda(factor(Status) ~  poly(smoking,arr[1],raw=T) + poly(cholesterol,arr[2],raw=T) +  poly(age,arr[3],raw=T) + poly(assay,arr[4],raw=T), data=train) )
qda.pred = predict(qda.fit, valid)$class
  if (mean(qda.pred==valid$Status) > bestVal) {
    bestArr = arr
    bestVal = mean(qda.pred==valid$Status)
  }
}
bestArr
bestVal
```


## GAMS

Best so far uses: factor(Status) ~  smoking + cholesterol + age +  assay

Various natural splines
```{r}
library(splines)
set.seed(1)
bestVal = 0
for (x in 1:500) {
  arr = sample(1:5, 4, replace=TRUE)
  gam.val = glm(factor(Status)~ns(smoking,arr[1])+ns(cholesterol,arr[2])+ns(age,arr[3])+ns(assay,arr[4]),data=train, family="binomial")
  gam.prob = predict(gam.val,valid)
  gam.prob = exp(gam.prob) / ( 1 + exp(gam.prob))
  gam.pred=rep(0,length(valid$Status))
  gam.pred[gam.prob>0.5]=1
  if (mean(gam.pred==valid$Status) > bestVal) {
    bestArr = arr
    bestVal = mean(gam.pred==valid$Status)
  }
}
bestArr
bestVal
```

Various smoothing splines
```{r}
library(gam)
set.seed(1)
bestVal = 0
par(mfrow =c(4,5))
for (x in 1:500) {
  arr = sample(1:5, 4, replace=TRUE)
  gam.val = gam(factor(Status)~s(smoking,arr[1])+s(cholesterol,arr[2])+s(age,arr[3])+s(assay,arr[4]),data=train, family="binomial")
  gam.prob = predict(gam.val,valid)
  gam.prob = exp(gam.prob) / ( 1 + exp(gam.prob))
  gam.pred=rep(0,length(valid$Status))
  gam.pred[gam.prob>0.5]=1
  table(gam.pred,valid$Status)
  if (mean(gam.pred==valid$Status) > bestVal) {
    bestArr = arr
    bestVal = mean(gam.pred==valid$Status)
  }
}
bestArr
bestVal
```
  
  
## TREE

```{r}
library(tree)
set.seed(1)
tree.info = tree(as.factor(Status)~.-Id,data=train)
summary(tree.info)
tree.prob=predict(tree.info,valid)
tree.pred=rep(0,length(valid$Status))
tree.pred[tree.prob[,2]>.5]=1
table(tree.pred,valid$Status)
mean(glm.pred==valid$Status)
plot(tree.info, main="Decision Tree")
text(tree.info,pretty=0,cex=0.4,col="Red")
```

Perform tree cross-validation
```{r}
tree.cv = cv.tree(tree.info,FUN=prune.misclass)
plot(tree.cv$size, tree.cv$dev, xlab="Tree Size",ylab="Dev", type="b", main="Tree Cross-Validation")
```

## Linear SMV

```{r}
library(e1071)
set.seed(1)
costArr = seq(0.001,0.03,0.0001)
tune.out=tune(svm,as.factor(Status)~.,data=train,kernel ="linear",
ranges =list(cost=costArr))
bestmod = tune.out$best.model
tempData = valid
tempData$Status = as.factor(tempData$Status)
yhat = predict(bestmod,tempData)
mean(yhat==valid$Status)
bestmod
```

## Polynomial SMV
```{r}
set.seed(1)
costArr = 10^seq(-2,2,0.3)
tune.out=tune(svm,as.factor(Status)~.,data=train,kernel ="polynomial",
ranges =list(cost=costArr, degree=seq(2:10)))
bestmod = tune.out$best.model
tempData = valid
tempData$Status = as.factor(tempData$Status)
yhat = predict(bestmod,tempData)
mean(yhat==valid$Status)
bestmod
```

## Radial SMV
```{r}
set.seed(1)
costArr = 10^seq(-2,2,0.3)
gammaArr = costArr
tune.out=tune(svm,as.factor(Status)~.,data=train,kernel ="radial",
ranges =list(cost=costArr, gamma=gammaArr))
bestmod = tune.out$best.model
tempData = valid
tempData$Status = as.factor(tempData$Status)
yhat = predict(bestmod,tempData)
mean(yhat==valid$Status)
bestmod
```